---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Denys Sharpylo"
  email: "sharpyld@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---
20/20. Good job

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
table(Cars93$Type)
round(table(Cars93$Type)/nrow(Cars93),3)
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
barplot(table(Cars93$Type), ylab = "Count", xlab = "Type", main = "Types of Cars", cex.axis = 1.2, cex.main=1.2, cex.names = 1.2, col =  c(1,3,5,7,8,12))
```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie(table(Cars93$Type), main = "Types of Cars in '93", col = c(1,3,5,7,8,12), cex=1.2 )
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.
```{r} 
table(Cars93$Type)
## Since "Midsize" type has the most occurences in the dataset(22), it is the mode of the variable "Type"
```

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hist(Cars93$Horsepower, cex.lab=1.2, cex.axis= 1.2, ylab = "Frequency", xlab = "Horsepower", main = "Cars93 Horsepower", col = c(4))
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hist(Cars93$Horsepower, cex.lab=1.2, cex.axis= 1.2, ylab = "Frequency", xlab = "Horsepower", main = "Cars93 Horsepower", col = c(4), breaks = seq(50,300,25))
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
```{r}
##Since there are a total of 93 entries in the dataset, the suggested rule of thumb is sqrt(93)= ~9-10. Therefore, the histogram with 25HP increments has a better number of classes (10). 
```


## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
```{r}
mode(Cars93$Passengers)
## Passengers is a discrete numerical variable. 
hist(Cars93$Passengers)
table(Cars93$Passengers)
## The problem with the graph is that there are columns that have no frequency/entries (e.g, 3 passengers in a car), but it's shown as if there are entries with 3-passenger cars. 
```

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
barplot(table(factor(Cars93$Passengers, levels = 1:10)), ylab = "Frequency", xlab = "# of Passengers", main = "Passengers in Cars93", cex.axis = 1.2, cex.names = 1.2, col = c(2,3,4,5,6,7))
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
hist(Cars93$Length, ylab = "Frequency", xlab = "Length", main = "Length of Cars93", cex.axis= 1.2, cex.lab=1.2, col = c(4))
```

b) What is the shape of the distribution?
```{r}
## approximately symmetrical
```

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
```{r}
mean(Cars93$Length)
median(Cars93$Length)
## In a symmetrical fully symmetrical graph, mean is equal to the median. However, since this graph is approximately symmetrical, we can see the deviation in the values of the mean and the median, though it is a small difference indicating how close to being symmetrical the above graph is.
```

You are done! 🏁 Don't forget to commit and push your .Rmd file to your Github repository before the due date.


