---
title: "Introduction to Supervised Learning <br>Part 1: Linear Regression"
author: "Eren Bilen <br> Dickinson College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(base_color = "#1c5253",colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
))

```


#Supervised versus Unsupervised Learning

**supervised learning** means that for each predictor $x_i$, $i=1,\ldots,n$ there is an associated response variable $y_i$, 

Our goal is to fit a model relating the response $(y)$ to the predictors $(x)$ such that we can accurately predict future responses (prediction) or such that we can better understand the relationship between the response and the predictor (inference). 

Examples of this include linear regression, logistic regression, boosting, support vector machines, regression tree.

---
#Supervised versus Unsupervised Learning

Unsupervised learning explores a different situation where for every observation $x_i$
that we observe, we do not observe a response variable $y_i$. 

Therefore, we cannot fit a model since we don't have a response variable! (All variales are $x_i$!)

In this setting, we lack the supervision of a response $y_i$ to guide the $x_i$ for model fitting, so we develop other methods to guide our analysis.

Typically we are **guided by the data**! 

Recall: k-means clustering, hierarchical clustering, LDA. 

The goal of clustering is to ascertain, on the basis of $x_1, \ldots, x_n$, where the observations fall into relatively distinct groups.

---
#Why Linear Regression

- Linear Regression is a simple approach for supervised learning and predictive and quantitative response. 
- It is one of the simplest methods to consider.
- Extremely widely used in research: Econometrics!
- Enables prediction + interpretation
  - Unlike tree based models or neural-nets where interpretation is almost impossible.


---
#Recall the Advertising data set

Consider Advertising data, where one is asked to suggest a marketing plan for next year that will result in high product sales.

Questions that we might want to address:

1. Is there a relationship between advertising budget and sales?
  * How strong is the relationship between advertising budget and sales?
2. Which media contribute to sales? (newspaper, TV, Superbowl..)
  * How accurately can we estimate the effect of each medium on sales? 
3. How accurately can we predict future sales, given all predictors?
4. What is the type of relationship? (linear, non-linear?)

Linear regression can answer each of these questions.

---
#Simple Linear Regression (SLR)

- SLR is a way for predicting a response $Y$ on the basis of a single predictor variable $X.$
 - $Y$ could be numeric, binary/categorical (coded as integers)
- Easily expandable to Multiple Linear Regression where there exist multiple predictors $X_i$.

We  model a linear relationship as 
$$\begin{equation}
  \hat{y} = \hat{\beta_o} + \hat{\beta_1} x,
\tag{1}
\end{equation}$$

- We often say that we are regressing $Y$ onto $X.$

In equation **(1)**, $\hat{\beta_o}$ and  $\hat{\beta_1}$ represent the slope and the intercept in the linear model. 

Once we have used our **training data** to produce estimates $\hat{\beta_o}, \hat{\beta_1}$ for the model coefficients, we can predict future sales on the basis of a particular value of TV advertising by computing. More on this to come.

---
#Estimating the Coefficients

We can use equation (1) to make predictions, but we need to "fit the model" and estimate the coefficients. 

Start with exploring the Advertising data


```{r,fig.height=3.75, fig.width=10}
ad <- read.table("data/Advertising.csv", header=TRUE, sep=",")
par(mfrow=c(1,3))
plot(ad$TV, ad$Sales, xlab="TV", ylab="Sales",cex=2,cex.lab=1.25)
plot(ad$Radio, ad$Sales, xlab="Radio", ylab="Sales",cex=2,cex.lab=1.25)
plot(ad$Newspaper, ad$Sales, xlab="Newspaper", ylab="Sales",cex=2,cex.lab=1.25)
```


---
#Estimating the Coefficients

- Let $\hat{y}_i = \hat{\beta_o} + \hat{\beta_1} x_i$ be the prediction $Y$ based on the $i$the value of X.
  - Goal: estimate the $\hat{\beta_o}$ (aka intercept) and $\hat{\beta_1}$ (aka slope) such that the resulting line goes through "right in the middle" of the data points.

- Define $e_i = y_i - \hat{y}_i$ known as the $i$th **residual**, or the **error** where $y_i$ is the actual value of the response variable, $\hat{y_i}$ is the model's prediction. 

- The **Sum of Squared Residiuals (SSR)** is defined as

\begin{align}
\text{RSS} &= e_1^2 + \ldots e_n^2 \\
&= \sum_{i=1}^n (y_i - \hat{\beta}_o \hat{\beta}_1x_i)^2
\end{align}

- Using calculus, one can show that the minimizers are 

\begin{align}
\hat{\beta}_1 &= \frac{\sum_{i=1}^n(x_i - \bar{x})(y_i - \bar{y})}{\sum_{i=1}^n(x_i - \bar{x})}, \; \hat{\beta}_0 = \bar{y} - \hat{\beta}_1 \bar{x}
\end{align}

---
# $R^2$

  $R^2$ is the **proportion of variance explained**—and it always takes on a value between 0 and 1, independent of the scale of Y.

$$R^2 = \frac{TSS-SSR}{TSS} = 1 -\frac{SSR}{TSS},$$
  where $TSS = \sum_i (y_i -\bar{y})$
  is the **total sum of squares**.

  - TSS measures the total variance in the response Y , and can be thought of as the amount of variability inherent in the response before the regression is performed.

- SSR measures the amount of variability that is left unexplained after performing the regression.

- $R^2$ measures the proportion of variability in Y that can be explained using X.

  - An $R^2$ statistic that is close to 1 indicates that a large proportion of the variability in the response has been explained by the regression.
  - A number near 0 indicates that the regression did not explain much of the variability in the response.
  - This might occur because the linear model is wrong, or the inherent variability in response is high, or both.
  - Additional benefit of regression: coefficients have meaning! (prediction vs. interpretation)

---

# R example

Load the MASS and ISLR packages
```{r,warning=FALSE}
library(MASS)
library(ISLR)
data(Boston)
```

We investivate the `Boston` data set, which records `medv` (median house value) for 506 neighborhoods around Boston.

**Goal**: predict `medv` using 13 predictors such as `rm` (average number of rooms per house), `age` (average age of houses), and `lstat` (percent of households with low socioeconomic status).

---
#Boston data set

```{r}
# check variable names
names(Boston)
# attach the Boston data set
attach(Boston)
```

What types of exploratory data
analysis can we do?
- histogram
- barplot
- scatterplot
- a map of Boston colored with a scale of average home price by neighborhood.

---
#Boston data set


- How do we do linear regression?

```{r}
# run the linear regression
lm.fit <- lm(medv~lstat ,data=Boston)
# limited standard output
lm.fit
```

- where `medv` is the Y-variable, `lstat` is the X variable.
- interpretation: one unit change in X is associated with a 0.95 *decrease* in Y.
- Does the interpretation make sense for `medv` and `lstat`?
  - Always check the mean Y and X when you interpret coefficients to get a sense of their magnitude.
---
#Boston data set

```{r}
# detailed summary output
summary(lm.fit)
```

---
# Regression plot

```{r,fig.width=5.5, fig.height=5.5}
# scatterplot
plot(lstat, medv)
# add the regression line
abline(lm.fit,lwd=2,col="red")
```

---
# Now, let's really talk about prediction

What does the model predict for the median home value in a neighborhood with `lstat`=10?

In supervised learning, you want to split your data into traning and test groups. This is done to prevent *overfitting*. 


```{r}
# create a training sample
trainingloc <- sample(1:nrow(Boston), 0.8*nrow(Boston))
training <- Boston[trainingloc, ]
# create a test sample
testloc <- setdiff(1:nrow(Boston), trainingloc) # setdiff finds the rows not in trainingloc
test <- Boston[testloc, ]

# fit the model
trained_model <- lm(medv~lstat ,data=training)

# predict for a single value
#trained_model$coefficients[1] + trained_model$coefficients[2] * 10

# equivalently,
#predict(trained_model, newdata = data.frame(lstat=c(10)))
```

---
# Now, let's really talk about prediction
```{r,fig.width=5,fig.height=5}
# predict for all lstat in the test sample, compare with the actual Y
preds <- trained_model$coefficients[1] + trained_model$coefficients[2] * test$lstat

# plot predicted (red) and actual medv (blue)
plot(test$lstat, preds,col=c('red'),ylab='medv', xlab='lstat',ylim=c(0,50),cex=1.25)
par(new=TRUE)
plot(test$lstat, test$medv,col=c('blue'),ann=FALSE, axes=FALSE,cex=1.25)
```

---
# How good is the model?: Mean Squared Error (MSE)
```{r}
errorsq <- (test$medv - preds)^2
mse <- sum(errorsq)/nrow(test)
mse
```

---
# Can we improve on the model? Add another variable
```{r}
# fit the model with the average number of rooms per house included as an additional predictor
trained_model <- lm(medv ~ lstat + rm ,data=training)

# predict for all lstat in the test sample, compare with the actual Y
preds <- trained_model$coefficients[1] + trained_model$coefficients[2] * test$lstat +
  trained_model$coefficients[3] * test$rm

# Get MSE
errorsq <- (test$medv - preds)^2
mse <- sum(errorsq)/nrow(test)
mse
```


---
# Works with binary outcome too!: LPM
```{r,fig.width=4,fig.height=4}
# fit the model with the average number of rooms per house included as an additional predictor
trained_model <- lm(chas ~ rm ,data=training)

# predict for all lstat in the test sample, compare with the actual Y
preds <- trained_model$coefficients[1] + trained_model$coefficients[2] * test$rm

# plot predicted (red) and actual medv (blue)
plot(test$rm, preds,col=c('red'),ylab='chas', xlab='rm',cex=1.25)
par(new=TRUE)
plot(test$rm, test$chas,col=c('blue'),ann=FALSE, axes=FALSE,cex=1.25)
```

---
# Works with binary outcome too!: LPM
```{r}
summary(trained_model)
```

---
# Works with binary outcome too!: LPM
```{r,warning=FALSE,message=FALSE}
library(dplyr)
test <- test %>% mutate(make_pred=(preds>mean(chas))*1)

# Get "confusion matrix"
oneone <- test %>% filter(chas==1, make_pred==1) %>% count()
onezero <- test %>% filter(chas==1, make_pred==0) %>% count()
zeroone <- test %>% filter(chas==0, make_pred==1) %>% count()
zerozero <- test %>% filter(chas==0, make_pred==0) %>% count()

percent_correct <- (oneone + zerozero)/nrow(test)
percent_correct
```



